### 2015-11-23 paper.tex
Initializing LaTeX paper

\documentclass[journal, a4paper]{IEEEtran}

\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

\usepackage{listings}

\usepackage{fancyvrb}

% some very useful LaTeX packages include:

%\usepackage{cite}      % Written by Donald Arseneau
                        % V1.6 and later of IEEEtran pre-defines the format
                        % of the cite.sty package \cite{} output to follow
                        % that of IEEE. Loading the cite package will
                        % result in citation numbers being automatically
                        % sorted and properly "ranged". i.e.,
                        % [1], [9], [2], [7], [5], [6]
                        % (without using cite.sty)
                        % will become:
                        % [1], [2], [5]--[7], [9] (using cite.sty)
                        % cite.sty's \cite will automatically add leading
                        % space, if needed. Use cite.sty's noadjust option
                        % (cite.sty V3.8 and later) if you want to turn this
                        % off. cite.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/

\usepackage{graphicx}   % Written by David Carlisle and Sebastian Rahtz
                        % Required if you want graphics, photos, etc.
                        % graphicx.sty is already installed on most LaTeX
                        % systems. The latest version and documentation can
                        % be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
                        % Another good source of documentation is "Using
                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
                        % which can be found as esplatex.ps and epslatex.pdf
                        % at: http://www.ctan.org/tex-archive/info/

%\usepackage{psfrag}    % Written by Craig Barratt, Michael C. Grant,
                        % and David Carlisle
                        % This package allows you to substitute LaTeX
                        % commands for text in imported EPS graphic files.
                        % In this way, LaTeX symbols can be placed into
                        % graphics that have been generated by other
                        % applications. You must use latex->dvips->ps2pdf
                        % workflow (not direct pdf output from pdflatex) if
                        % you wish to use this capability because it works
                        % via some PostScript tricks. Alternatively, the
                        % graphics could be processed as separate files via
                        % psfrag and dvips, then converted to PDF for
                        % inclusion in the main file which uses pdflatex.
                        % Docs are in "The PSfrag System" by Michael C. Grant
                        % and David Carlisle. There is also some information
                        % about using psfrag in "Using Imported Graphics in
                        % LaTeX2e" by Keith Reckdahl which documents the
                        % graphicx package (see above). The psfrag package
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/psfrag/

%\usepackage{subfigure} % Written by Steven Douglas Cochran
                        % This package makes it easy to put subfigures
                        % in your figures. i.e., "figure 1a and 1b"
                        % Docs are in "Using Imported Graphics in LaTeX2e"
                        % by Keith Reckdahl which also documents the graphicx
                        % package (see above). subfigure.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/subfigure/

\usepackage{url}        % Written by Donald Arseneau
                        % Provides better support for handling and breaking
                        % URLs. url.sty is already installed on most LaTeX
                        % systems. The latest version can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/other/misc/
                        % Read the url.sty source comments for usage information.

%\usepackage{stfloats}  % Written by Sigitas Tolusis
                        % Gives LaTeX2e the ability to do double column
                        % floats at the bottom of the page as well as the top.
                        % (e.g., "\begin{figure*}[!b]" is not normally
                        % possible in LaTeX2e). This is an invasive package
                        % which rewrites many portions of the LaTeX2e output
                        % routines. It may not work with other packages that
                        % modify the LaTeX2e output routine and/or with other
                        % versions of LaTeX. The latest version and
                        % documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/sttools/
                        % Documentation is contained in the stfloats.sty
                        % comments as well as in the presfull.pdf file.
                        % Do not use the stfloats baselinefloat ability as
                        % IEEE does not allow \baselineskip to stretch.
                        % Authors submitting work to the IEEE should note
                        % that IEEE rarely uses double column equations and
                        % that authors should try to avoid such use.
                        % Do not be tempted to use the cuted.sty or
                        % midfloat.sty package (by the same author) as IEEE
                        % does not format its papers in such ways.

\usepackage{amsmath}    % From the American Mathematical Society
                        % A popular package that provides many helpful commands
                        % for dealing with mathematics. Note that the AMSmath
                        % package sets \interdisplaylinepenalty to 10000 thus
                        % preventing page breaks from occurring within multiline
                        % equations. Use:
%\interdisplaylinepenalty=2500
                        % after loading amsmath to restore such page breaks
                        % as IEEEtran.cls normally does. amsmath.sty is already
                        % installed on most LaTeX systems. The latest version
                        % and documentation can be obtained at:
                        % http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% Other popular packages for formatting tables and equations include:

%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty which improves the
% LaTeX2e array and tabular environments to provide better appearances and
% additional user controls. array.sty is already installed on most systems.
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% V1.6 of IEEEtran contains the IEEEeqnarray family of commands that can
% be used to generate multiline equations as well as matrices, tables, etc.

% Also of notable interest:
% Scott Pakin's eqparbox package for creating (automatically sized) equal
% width boxes. Available:
% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/eqparbox/

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.

% Your document starts here!
\begin{document}

% Define document title and author
	\title{Trustability of certificates}
	\author{Alexandre Kervadec
	\thanks{Tutor : A.Guermouche}}
	\markboth{University of Bordeaux - Master 2 Computing Sciences}{}
	\maketitle

% Write abstract here
\begin{abstract}

\end{abstract}

% Each section begins with a \section{title} command
\section{}

% Your document ends here!
\end{document}

### 2015-11-29 paper.tex
First raw final paper

@@
 % Each section begins with a \section{title} command
-\section{}
+\section{SKIP, X.509, CAs and PGP}
+
+\subsection{SKIP (Simple Key-Management for Internet Protocol)}
+
+[...]
+
+\subsection{X.509 and CAs}
+
+Description of the different entities
+
+1. CA : can be public (like banks with clients), commercial (like Verisign) or private (like internal departement of a compagny, to log user)
+2. Subscriber : sends some infos to the CA to add it to his certificate
+3. User : ask infos to CA(s), it's central to the process, since the user party is relying on the informations and is thus at risk
+
+Concerns about the authentification services provided by CAs
+
+1. The content of a certificate needs to be discussed, as well as certificate revocation
+2. Issues about DN (Distinguished Name) and CA : a user can possess one or more DN, and use one or more DN on one or more DN
+3. Validation of the user, using an ID, which is easily subject to fraud
+
+Going deeper with user validation : DN scheme based on X.500 Recomendation, but it is not completly defined, and will (in 1998) probably not be.
+Also, X509 certification depends on many others such as ISO, ANSI, ITU and IETF.
+Thus lead to a lack of harmonization.
+Plus, there is a big problem with CPS (Certification Practice Statements), that also can be seen such as flexibility (for pros), because each CA answer specific needs, so no harmonization again.
+
+Some kind of conclusion about harmonization (lack of), in a world wide vision.
+
+\subsection{PGP}
+
+PGP \footnote{Pretty Good Privacy}, globaly created by *put his name*.
+
+%&\include
+[Schema]
+It is based on the *web-of-trust*, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
+In that ring, even if the client doesn't know everyone, someone will know him, thus the web-of-trust principle.
+Each member of the ring have a trust evaluation of CAs. If the client knows someone he trust that trust the CA he doesn't know, then he accept the certificate and trust the CA.
+But if the client trust someone who doesn't trust the CA, he reject the CA.
+The big point is when nobody knows the CA, and when trusted friends of the client have a medium evaluation of the CA : when to trust, when to reject, at which level of trust?
 

### 2015-11-30 paper.tex
paper in progress

@@
 \begin{abstract}
-
+This paper is inspired from E.Gerck "Overview of Certification systems: X.509, CA, PGP and SKIP" \cite{gerck1998overview}
 \end{abstract}
@@
 
-PGP \footnote{Pretty Good Privacy}, globaly created by *put his name*.
+PGP \footnote{Pretty Good Privacy}, created thanks to Phil Zimmermann researches.
 
@@
 [Schema]
-It is based on the *web-of-trust*, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
+It is based on the \textit{web-of-trust}, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
 In that ring, even if the client doesn't know everyone, someone will know him, thus the web-of-trust principle.
@@
 
+\nocite{*}
+
+\begin{thebibliography}{99}
+
+\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998, http://www.blackhat.com/presentations/bh-usa-99/EdGerck/certover.pdf
+
+\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010, http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html
+
+\end{thebibliography}
+
 % Your document ends here!

### 2015-12-09 paper.tex
progress on overview and reshaping plan

@@
 % Define document title and author
-	\title{Trustability of certificates}
+	\title{Trustability of certification systems}
 	\author{Alexandre Kervadec
@@
 \begin{abstract}
-This paper is inspired from E.Gerck "Overview of Certification systems: X.509, CA, PGP and SKIP" \cite{gerck1998overview}
+This paper is inspired from E.Gerck "Overview of Certification systems: X.509, CA, PGP and SKIP" \cite{gerck1998overview}.
+It gives an overview of (main of) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google.
+Then it gives a reflection on pros and cons of each system, on a technical point of view and about the government stranglehold on data exchanges.
 \end{abstract}
@@
 % Each section begins with a \section{title} command
-\section{SKIP, X.509, CAs and PGP}
+\section{Overview of certification systems}
 
@@
 
-%&\include
-[Schema]
+[TODO : Schema]
+
 It is based on the \textit{web-of-trust}, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
@@
 The big point is when nobody knows the CA, and when trusted friends of the client have a medium evaluation of the CA : when to trust, when to reject, at which level of trust?
+
+\subsection{DANE}
+
+DANE\footnote{DNS-Based Authentication of Named Entities} is not implemented yet, but the IETF is working on the standard.
+
+\subsection{Cetificate Transparency}
+
+Because of a case of Google CA's usurpation, Google is working on a new system that can improve certification systems.
 
@@
 
+\bibitem{dane2012} P Hoffman, J Schlyter, The DNS-based authentication of named entities (DANE) transport layer security (TLS) protocol: TLSA, 2012, https://www.rfc-editor.org/rfc/pdfrfc/rfc6698.txt.pdf
+
 \end{thebibliography}

### 2015-12-09 paper.tex
progressing on redaction

--- 2015-11-29 paper.tex
@@
 \begin{abstract}
-
+This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
+It first gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with a reflection on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
 \end{abstract}
@@
 
-Description of the different entities
+X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.
+With this kind of certification system, there are three different entities, which are :
 
-1. CA : can be public (like banks with clients), commercial (like Verisign) or private (like internal departement of a compagny, to log user)
-2. Subscriber : sends some infos to the CA to add it to his certificate
-3. User : ask infos to CA(s), it's central to the process, since the user party is relying on the informations and is thus at risk
+\begin{enumerate}
+	\item \textbf{CA : Certification Authority}, an entity at controls the authentication services and the management of certificates.
+It can be public (like banks with their clients), commercial (like Verisign which sells its services) or private (like a compagny department, for an internal purpose).
+	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. The entity is one that need to be trusted by the next entity (\textit{user}).
+	\item \textbf{User :} ask infos to CA(s), it's central to the process, since the user party is relying on the informations and is thus at risk.
+\end{enumerate}
 
-Concerns about the authentification services provided by CAs
+The main concerns about authentification services provided by CAs are :
 
-1. The content of a certificate needs to be discussed, as well as certificate revocation
-2. Issues about DN (Distinguished Name) and CA : a user can possess one or more DN, and use one or more DN on one or more DN
-3. Validation of the user, using an ID, which is easily subject to fraud
+\begin{itemize}
+	\item The content of a certificate needs to be discussed, as well as certificate revocation
+	\item Issues about DN (Distinguished Name) and CA : a user can possess one or more DN, and use one or more DN on one or more DN
+	\item Validation of the user, using an ID, which is easily subject to fraud
+\end{itemize}
 
@@
 
-PGP \footnote{Pretty Good Privacy}, globaly created by *put his name*.
+PGP\footnote{Pretty Good Privacy}, created thanks to Phil Zimmermann researches.
 
-%&\include
-[Schema]
-It is based on the *web-of-trust*, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
+[TODO : Schema]
+
+It is based on the \textit{web-of-trust}, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
 In that ring, even if the client doesn't know everyone, someone will know him, thus the web-of-trust principle.
@@
 
+\subsection{DANE}
+
+DANE\footnote{DNS-Based Authentication of Named Entities} is not implemented yet, but a IETF team is working on the standard.
+
+\subsection{Cetificate Transparency}
+
+Because of a case of Google CA's usurpation, Google is working on a new system that can improve certification systems.
+
+\nocite{*}
+
+\begin{thebibliography}{99}
+
+\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998. http://www.blackhat.com/presentations/bh-usa-99/EdGerck/certover.pdf
+
+\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010. http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html
+
+\bibitem{dane2012} P Hoffman, J Schlyter, The DNS-based authentication of named entities (DANE) transport layer security (TLS) protocol: TLSA, 2012.
+https://www.rfc-editor.org/rfc/pdfrfc/rfc6698.txt.pdf
+
+\bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf
+
+\bibitem{rfc3647} Solo, D., Housley, R., \& Ford, W. (1999). Internet X. 509 public key infrastructure certificate and CRL profile. http://tools.ietf.org/html/rfc2459
+
+\end{thebibliography}
+
 % Your document ends here!

### 2015-12-10 paper.tex
progress...

@@
 This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
-It first gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with a reflection on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
+It first gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
 \end{abstract}
@@
 
-\subsection{SKIP (Simple Key-Management for Internet Protocol)}
-
-[...]
+First sections talk about existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
+Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of certification that may answer to current issues.
 
 \subsection{X.509 and CAs}
+\label{x509}
 
@@
 
-\subsection{PGP}
+\subsection{PGP (Pretty Good Privacy)}
+\label{pgp}
 
-PGP\footnote{Pretty Good Privacy}, created thanks to Phil Zimmermann researches.
+PGP, created thanks to Phil Zimmermann researches.
 
@@
 
-\subsection{DANE}
+\subsection{SKIP (Simple Key-Management for Internet Protocol)}
+\label{skip}
 
-DANE\footnote{DNS-Based Authentication of Named Entities} is not implemented yet, but a IETF team is working on the standard.
+\subsection{DANE (DNS-Based Authentication of Named Entities)}
+\label{dane}
+
+DANE is not implemented yet, but a IETF team is working on the standard.
 
 \subsection{Cetificate Transparency}
+\label{certtrans}
 
@@
 \bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998. http://www.blackhat.com/presentations/bh-usa-99/EdGerck/certover.pdf
+
+\bibitem{skip1995} Ashar Aziz, Tom Markson, Hemma Prafullchandra, Sun Microsystems, Inc., v.06, 1995. https://tools.ietf.org/html/draft-ietf-ipsec-skip-06
 

### 2015-12-22 paper.tex
X.509 & PGP done

@@
 \usepackage{fancyvrb}
-
-% some very useful LaTeX packages include:
-
-%\usepackage{cite}      % Written by Donald Arseneau
-                        % V1.6 and later of IEEEtran pre-defines the format
-                        % of the cite.sty package \cite{} output to follow
-                        % that of IEEE. Loading the cite package will
-                        % result in citation numbers being automatically
-                        % sorted and properly "ranged". i.e.,
-                        % [1], [9], [2], [7], [5], [6]
-                        % (without using cite.sty)
-                        % will become:
-                        % [1], [2], [5]--[7], [9] (using cite.sty)
-                        % cite.sty's \cite will automatically add leading
-                        % space, if needed. Use cite.sty's noadjust option
-                        % (cite.sty V3.8 and later) if you want to turn this
-                        % off. cite.sty is already installed on most LaTeX
-                        % systems. The latest version can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/cite/
-
-\usepackage{graphicx}   % Written by David Carlisle and Sebastian Rahtz
-                        % Required if you want graphics, photos, etc.
-                        % graphicx.sty is already installed on most LaTeX
-                        % systems. The latest version and documentation can
-                        % be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-                        % Another good source of documentation is "Using
-                        % Imported Graphics in LaTeX2e" by Keith Reckdahl
-                        % which can be found as esplatex.ps and epslatex.pdf
-                        % at: http://www.ctan.org/tex-archive/info/
-
-%\usepackage{psfrag}    % Written by Craig Barratt, Michael C. Grant,
-                        % and David Carlisle
-                        % This package allows you to substitute LaTeX
-                        % commands for text in imported EPS graphic files.
-                        % In this way, LaTeX symbols can be placed into
-                        % graphics that have been generated by other
-                        % applications. You must use latex->dvips->ps2pdf
-                        % workflow (not direct pdf output from pdflatex) if
-                        % you wish to use this capability because it works
-                        % via some PostScript tricks. Alternatively, the
-                        % graphics could be processed as separate files via
-                        % psfrag and dvips, then converted to PDF for
-                        % inclusion in the main file which uses pdflatex.
-                        % Docs are in "The PSfrag System" by Michael C. Grant
-                        % and David Carlisle. There is also some information
-                        % about using psfrag in "Using Imported Graphics in
-                        % LaTeX2e" by Keith Reckdahl which documents the
-                        % graphicx package (see above). The psfrag package
-                        % and documentation can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/psfrag/
-
-%\usepackage{subfigure} % Written by Steven Douglas Cochran
-                        % This package makes it easy to put subfigures
-                        % in your figures. i.e., "figure 1a and 1b"
-                        % Docs are in "Using Imported Graphics in LaTeX2e"
-                        % by Keith Reckdahl which also documents the graphicx
-                        % package (see above). subfigure.sty is already
-                        % installed on most LaTeX systems. The latest version
-                        % and documentation can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/subfigure/
-
-\usepackage{url}        % Written by Donald Arseneau
-                        % Provides better support for handling and breaking
-                        % URLs. url.sty is already installed on most LaTeX
-                        % systems. The latest version can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/contrib/other/misc/
-                        % Read the url.sty source comments for usage information.
-
-%\usepackage{stfloats}  % Written by Sigitas Tolusis
-                        % Gives LaTeX2e the ability to do double column
-                        % floats at the bottom of the page as well as the top.
-                        % (e.g., "\begin{figure*}[!b]" is not normally
-                        % possible in LaTeX2e). This is an invasive package
-                        % which rewrites many portions of the LaTeX2e output
-                        % routines. It may not work with other packages that
-                        % modify the LaTeX2e output routine and/or with other
-                        % versions of LaTeX. The latest version and
-                        % documentation can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/contrib/supported/sttools/
-                        % Documentation is contained in the stfloats.sty
-                        % comments as well as in the presfull.pdf file.
-                        % Do not use the stfloats baselinefloat ability as
-                        % IEEE does not allow \baselineskip to stretch.
-                        % Authors submitting work to the IEEE should note
-                        % that IEEE rarely uses double column equations and
-                        % that authors should try to avoid such use.
-                        % Do not be tempted to use the cuted.sty or
-                        % midfloat.sty package (by the same author) as IEEE
-                        % does not format its papers in such ways.
-
-\usepackage{amsmath}    % From the American Mathematical Society
-                        % A popular package that provides many helpful commands
-                        % for dealing with mathematics. Note that the AMSmath
-                        % package sets \interdisplaylinepenalty to 10000 thus
-                        % preventing page breaks from occurring within multiline
-                        % equations. Use:
-%\interdisplaylinepenalty=2500
-                        % after loading amsmath to restore such page breaks
-                        % as IEEEtran.cls normally does. amsmath.sty is already
-                        % installed on most LaTeX systems. The latest version
-                        % and documentation can be obtained at:
-                        % http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% Other popular packages for formatting tables and equations include:
-
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty which improves the
-% LaTeX2e array and tabular environments to provide better appearances and
-% additional user controls. array.sty is already installed on most systems.
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-% V1.6 of IEEEtran contains the IEEEeqnarray family of commands that can
-% be used to generate multiline equations as well as matrices, tables, etc.
-
-% Also of notable interest:
-% Scott Pakin's eqparbox package for creating (automatically sized) equal
-% width boxes. Available:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/supported/eqparbox/
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+\usepackage{graphicx}
+\usepackage{url}
+\usepackage{amsmath}
+\usepackage{tikz}
+\usepackage{verbatim}
 
@@
 	\thanks{Tutor : A.Guermouche}}
-	\markboth{University of Bordeaux - Master 2 Computing Sciences}{}
+	\markboth{University of Bordeaux -- Master 2 Computing Sciences - Networking and mobile systems}{}
 	\maketitle
@@
 This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
-It first gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
+First, it gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
 \end{abstract}
@@
 First sections talk about existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
-Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of certification that may answer to current issues.
+Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of digital certification which may solve current issues.
 
@@
 
-X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.
+X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.\\
 With this kind of certification system, there are three different entities, which are :
@@
 \begin{enumerate}
-	\item \textbf{CA : Certification Authority}, an entity at controls the authentication services and the management of certificates.
-It can be public (like banks with their clients), commercial (like Verisign which sells its services) or private (like a compagny department, for an internal purpose).
-	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. The entity is one that need to be trusted by the next entity (\textit{user}).
-	\item \textbf{User :} ask infos to CA(s), it's central to the process, since the user party is relying on the informations and is thus at risk.
+	\item \textbf{CA : Certification Authority}, an entity that controls authentication services and management of digital certificates.
+It can be public (like banks with their clients), commercial (such as \textit{Verisign} which sells its services) or private (like a compagny department, for an internal purposes).
+	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. This entity is one that need to be trusted by the \textit{user} entity.
+	\item \textbf{User :} ask informations about \textit{subscribers} to CA(s), it's central to the process, since the user party is relying on the informations given by \textit{CAs} and is thus at risk.
 \end{enumerate}
+
+The way that certification system is the following:
+\begin{itemize}
+	\item A subscriber gives informations to the CA, in order to create a digital certificate for itself in the CA directory.
+	\item The CA receive those informations and put it in a certificate.
+	\item A user (the public in general) receive a certificate from a subscriber.
+In order to authenticate this subscriber, the client ask to the CA if the transmitted certificate matches with the subscriber's identity.
+	\item The CA answer to the user to give him the precious informations about the subscriber's identity.
+\end{itemize}
+
+There is a not outwardly perceived entity in this system, it is the Naming Authority (DN).
+It provides to the CA a unique distinguished name (DN) which matches to the subscriber identity in the certificate.\\
+Note that the CA can double in NA, but they provide two different services: the CA certificate refers to a name but does not denote it, the NA does.\\
 
@@
 \begin{itemize}
-	\item The content of a certificate needs to be discussed, as well as certificate revocation
-	\item Issues about DN (Distinguished Name) and CA : a user can possess one or more DN, and use one or more DN on one or more DN
-	\item Validation of the user, using an ID, which is easily subject to fraud
+	\item The content of a certificate needs to be discussed, as well as certificate revocation.
+For example, a subscriber can generate multiple DN for a same CA or the same DN to multiple CAs, so maybe, a better subscriber identificator need to be added.
+	\item Are the validation procedures for the certified data that is included in a certificate solid enough?
+Because each CA has its own self-defined rules (Certification Practice Statement\footnote{CPS}), which can by completly different from one CA to another.
 \end{itemize}
@@
 
-Some kind of conclusion about harmonization (lack of), in a world wide vision.
+Also, regarrding to the law, CAs deny all leak of information from themselves.
+
+Some kind of conclusion about harmonization (lack of), in a world wide vision, if space there is.
 
@@
 
-PGP, created thanks to Phil Zimmermann researches.
+PGP, created thanks to Phil Zimmermann researches. It has two parts: certification and encryption, following text only deals with certification.\\
+Compared to X.509, PGP is more \textit{internet-like}, this because of its \textit{introducer-model} base.\\
+PGP depends on a chain of authentificators: the users themselves.
+This chain forms a ring, but not in a closed way but in a mathematical way like a list or a \textit{web-of-trust}.
+You may not know everyone in this ring, but you can assume that you will know somebody who know this user.
+You can also assume that different rings can have some contact points to garantee the referrals.
 
-[TODO : Schema]
+\begin{figure}[!hbt]
+	\begin{center}
 
-It is based on the \textit{web-of-trust}, if the client doesn't know a CA which gave him a certificate, he ask to the PGP ring.
-In that ring, even if the client doesn't know everyone, someone will know him, thus the web-of-trust principle.
-Each member of the ring have a trust evaluation of CAs. If the client knows someone he trust that trust the CA he doesn't know, then he accept the certificate and trust the CA.
-But if the client trust someone who doesn't trust the CA, he reject the CA.
-The big point is when nobody knows the CA, and when trusted friends of the client have a medium evaluation of the CA : when to trust, when to reject, at which level of trust?
+		\begin{tikzpicture}[shorten >=1pt,->]
+			\tikzstyle{vertex}=[circle,fill=black!25,minimum size=25pt,inner sep=0pt]
+
+			\foreach \name/\angle/\text in {P-1/234/Alice, P-2/162/Bob, 
+                                  P-3/90/Oscar, P-4/18/Jack, P-5/-54/Alex}
+				\node[vertex,xshift=6cm,yshift=.5cm] (\name) at (\angle:2cm) {$\text$};
+
+			\draw (P-1) -- (P-5);
+			\draw[dashed,->] (P-2) -- (P-4);
+			\draw (P-4) -- (P-3);
+			\draw (P-3) -- (P-1);
+			\draw (P-1) -- (P-2);
+			\draw (P-2) -- (P-1);
+			\draw (P-3) -- (P-5);
+			\draw (P-5) -- (P-3);
+
+		\end{tikzpicture}
+		\caption{Communities of trust: Jack introduces Oscar to Bob's public-key certificate before Bob receives it.}
+		\label{fig:web-of-trust}
+	\end{center}
+\end{figure}
+
+Thus, we can figure out that PGP breaks the traditional hierarchical trust architecture with this \textit{web-of-trust} approach.\\
+
+Let's get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}}: Bob want to exchange datas with Oscar, which he didn't knew before.
+But Jack, a colleague of Bob's, signs Oscar's public-key certificate whch he knows is authentic.\\
+Oscar then forwards his signed certificate to Bob who wishes to communicate with Oscar privately.\\
+Bob, who knows and truts Jack as an \textit{introducer}, find out, after verification, that Jack is among Oscar's certificate signer.
+Therefor, Bob can be confident that Oscar's public key is authentic.
+
+However, had Bob not known or trusted Oscar's signers, including Jack, he would have been skeptical about the authenticity of Oscar's public-key.
+Oscar would have to find another introducer whom Bob trusts to sign Oscar's public-key certificate.
+This is illustrated in Figure \ref{fig:web-of-trust}.\\
+This is illustrated in Figure 1, where the dotted arrow indicates ''Bob trusts Jack as introducer'' and the solid arrow indicates ''Jack trusts Oscar’s public key validity''.
+
+PGP also includes a public-key certificate and an introducer trustwhorthiness. They have different levels, for a public-key certificate they are:
+\begin{itemize}
+	\item \textit{Undefined:} we cannot say wether this public-key is valid or not.
+	\item \textit{Marginal:} this public-key \textit{may} be valid but we cannot be sure.
+	\item \textit{Complete:} we can be fully confident that this public-key is valid.
+\end{itemize}
+For an introducer, those levels of trustworthiness are the following:
+\begin{itemize}
+	\item \textit{Full:} this public-key is fully trusted to introduce another public-key.
+	\item \textit{Marginal:} this public-key can be trusted to introduce another public-key, but, it is uncertain whether it is fully competent to do that.
+	\item \textit{Untrustworthy:} this public-key should not be trustedto introduce another, therefore any occurence of this key as a signature on another public-key should be ignored.
+\end{itemize}
+
+Only the trustworthiness of public-key's validity is automatically evaluated by PGP.
+Introducer's worthiness is manually assigned by each user of to the public key, and exists only within each individual user's public-key ring\footnote{PGP allows users to have files representing multipke \textit{key rings} to store public or secret keys\cite{caronni2000}.}.
+
+The big point is the use of such a tool in a world wide range: that is raisonnably not possible.
+We can assume to use it in a little group, where trustworthiness is effective, but in a more wide area, it is problematic to trust such a amount of people, or find trusted introducers.
 
@@
 
-\subsection{Cetificate Transparency}
+\subsection{CATA: Certificate Authority Transparency and Auditability}
 \label{certtrans}
@@
 
-\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998. http://www.blackhat.com/presentations/bh-usa-99/EdGerck/certover.pdf
+\bibitem{gerck1998overview} E.Gerck, Overview of Certification Systems: X. 509, CA, PGP and SKIP, 1998.
 
@@
 
-\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010. http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html
+\bibitem{skip2010user} Oracle, SunScreen SKIP User's Guide, Release 1.1, 2010. %http://docs.oracle.com/cd/E19957-01/805-5743/6j5dvnrfs/index.html
+
+\bibitem{caronni2000} Caronni, G.
+(2000).
+Walking the web of trust.
+In Enabling Technologies: Infrastructure for Collaborative Enterprises, 2000.(WET ICE 2000).
+Proeedings.
+IEEE 9th International Workshops on (pp.
+153-158).
+IEEE.
+%http://system.khl.mydns.jp/college/WOT/The_PGP_Trust_Model.pdf
 
@@
 
-\bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf
+\bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). %https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf
 
-\bibitem{rfc3647} Solo, D., Housley, R., \& Ford, W. (1999). Internet X. 509 public key infrastructure certificate and CRL profile. http://tools.ietf.org/html/rfc2459
+\bibitem{rfc3647} Solo, D., Housley, R., \& Ford, W. (1999). Internet X. 509 public key infrastructure certificate and CRL profile. %http://tools.ietf.org/html/rfc2459
 

### 2015-12-22 paper.tex
SKIP done

@@
 We can assume to use it in a little group, where trustworthiness is effective, but in a more wide area, it is problematic to trust such a amount of people, or find trusted introducers.
+Furthermore, in a commercial use, there is no entity responsible when something goes wrong.
 
@@
 \label{skip}
+
+SKIP implements a linked-chain of two-sided node authenticators, where each node derives its informations from a sort of \textit{directory service}.\\
+The problem is that SKIP happens at a low level, so it is transparent to the user; also assuming that the user has no control over security or other certification choices.
+So it has to be complemented by a second higher level protocol.
+Furthermore, SKIP does not support IP translation (such as NAT behind a firewall), that makes it quite useless.\\
+So that protocol imediatly looks unapropriate for our study. In other way, SKIP does not exist anymore, so it is pointless to go further in it study.
+
+\subsection{Thoughts about X.509 and PGP}
+\label{thoughts_old}
 

### 2015-12-22 paper.tex
Old tech done, maybe need to rewrite the end of X509

@@
 
-\subsection{Thoughts about X.509 and PGP}
+\subsection{Thoughts about X.509, CAs and PGP}
 \label{thoughts_old}
+
+To be used in a commercial way (for responsabilities purpose), X.509, CAs and PGP need a centralized certification control system.
+However, there is a struggle between this need of centralization and the internet architecture (that is decentralized).
+This show us the uselessness of centralized governement controls, since we have no central government on the internet which give a world governance.\\
+So the question emerging in there is: how can we (and do we need to) have a main control on an internet environement, to assure the goodness of datas and legitimacy of interlocutors?\\
+Some answers have been made since then (thus since the publication of the original article), answers that we are going to explain further.
 
@@
 
-Because of a case of Google CA's usurpation, Google is working on a new system that can improve certification systems.
+This protocol has been developped by a Google team few years ago \cite{LL11}.
+It is still in developpment \cite{certtransp2013} so it is not implemented and deployed yet.
+It also has to be improve, like everythign in computing sciences, in the early years.\\
+CATA's purpose is a public list of all certificates that are emitted by CAs.
+Thus, clients are allowed to search in that public list, and verify that the received certificate is in that list, to be validate (if it is not in the list, it shall be rejected).\\
+In that way, a(conscientious) domain administrator can have access to that list and monitor all fraudulent certificates for his domain.\\
+For the moment, there are plenty of questions to answer like :
+\begin{itemize}
+	\item Who should hold those lists?
+	\item How many lists should be created?
+	\item How does the revokation should work?
+	\item What should happen if a certificate is no longer available (path unreachable)?
+\end{itemize}
 
@@
 
+\bibitem{LL11} Ben Laurie and Adam Langley. Certificate authority transparency and auditability. http://www.links.
+org/files/CertificateAuthorityTransparencyandAuditability.pdf, 2011.
+
 \bibitem{certtransp2013} Laurie, B., Langley, A., \& Kasper, E. (2013). Certificate transparency (No. RFC 6962). %https://www.rfc-editor.org/rfc/pdfrfc/rfc6962.txt.pdf

### 2015-12-22 paper.tex
Begining DANE

@@
 
-DANE is not implemented yet, but a IETF team is working on the standard.
+DANE is not implemented yet, but a IETF team is working on the standard.\\
+It has begin with a simple assert: why use a new trust infrastructure if we already have to use DNS to resolve a domain name?\\
+From here, DNS must be securised, knowing that it is easily to set a MITM with DNS cache poisoning \cite{bor08} \cite{aa04}.
+That is the main purpose of DNSSEC: secure DNS.
+DANE allow to store keys in DNS.
+Thereby, \verb$www.example.org$ would be able to secure \verb$https://www.example.org$ without using a suspicious third part, just by putting its keys in the domain name that it controls.
+By its arborescent nature, \verb$example.org$ shall not be able to cheat on \verb$example.com$.
 
@@
 
+\bibitem{bor08} Stéphane Bortzmeyer. Comment fonctionne la faille kaminsky ? http://www.bortzmeyer.org/comment-fonctionne-la-faille-kaminsky.html, 2008.
+
+\bibitem{aa04} D. Atkins and R. Austein. Rfc 3833 : Threat analysis of the domain name system (dns). http://www.
+ietf.org/rfc/rfc3833.txt, 2004.
+
 \bibitem{dane2012} P Hoffman, J Schlyter, The DNS-based authentication of named entities (DANE) transport layer security (TLS) protocol: TLSA, 2012.

### 2015-12-24 paper.tex
DANE -> Done
Miss one point on DANE is the last kilometer faillure.

@@
 % Define document title and author
-	\title{Trustability of certification systems}
+	\title{From a centralized certification systems to a more internet-like one}
 	\author{Alexandre Kervadec
@@
 It has begin with a simple assert: why use a new trust infrastructure if we already have to use DNS to resolve a domain name?\\
-From here, DNS must be securised, knowing that it is easily to set a MITM with DNS cache poisoning \cite{bor08} \cite{aa04}.
+From here, DNS must be securised, knowing that it is easily to set a MITM with DNS cache poisoning \cite{aa04} using the Kaminsky attack \cite{bor08}.
 That is the main purpose of DNSSEC: secure DNS.
@@
 Thereby, \verb$www.example.org$ would be able to secure \verb$https://www.example.org$ without using a suspicious third part, just by putting its keys in the domain name that it controls.
-By its arborescent nature, \verb$example.org$ shall not be able to cheat on \verb$example.com$.
+By its arborescent nature, \verb$example.org$ shall not be able to cheat on \verb$example.com$.\\
+
+Lets give us an example:\\
+From the classic schem, we have \textit{Alice}, the subscriber (\verb$alice.example.com$), \textit{Bob} its client, and \textit{Charlie}, the serious AC. There are three different scenarios:
+\begin{itemize}
+	\item \textit{Type 0:} Alice has got a certificate delivered by Charlie and is really satisfied by it.
+But Alice is afraid that \textit{Oscar} (a bad AC) forge a false certificate for its website.
+So Alice uses DANE to tell in DNS that Charlie and only Charlie is its AC.
+	\item \textit{Type 1:} Alice has got a certificate delivered by Charlie, but fear that Charlie give its certificate to bad people. So Alice uses DANE to publish in DNS the origin of its certificate.
+	\item \textit{Type 2:} that is the \textit{max scenario} in DANE. Alice doesn't trust ACs anymore, so it decides to emite it via DANE on DNS, that gives the entire responsability on DNS servers. 
+\end{itemize}
+
+This type is in a new register called TLSA\footnote{TLSA doesn't stand for anything: it is just the name of the RRtype (Record Resource type).} in DNS. Her is an example of a TLSA register:\\
+\begin{Verbatim}[fontsize=\small]
+_443._tcp.www.example.com. IN TLSA (
+	0 0 1 d2abde240d7cd3ee6b4b28c54df0&34b9
+	7983a1d16e8a410e4561cb106618e971 )
+\end{Verbatim}
+With:
+\begin{itemize}
+	\item \verb$443$ : the port used.
+	\item \verb$tcp$ : the protocole used.
+	\item \verb$www.example.com$ the domain name.
+	\item \verb$0 x x$ : the type used, here it is a type 0 (\verb$1 x x$ for type 1 and \verb$2 x x$ for type 2).
+	\item The rest is the certificate itself.
+\end{itemize}
+
+As we said before, DNS is not secured, so DNSEC has been deployed.
+DNSSEC is a protocol that allow to the DNS server (master) ti encrypt and sign its answer with a public key cryptographic system.
+It also permit to return a \verb$NXDOMAIN$ type (''No Such Domain'').\\
+Nervertheless DNSSEC has some vulnaribilities such as:
+\begin{itemize}
+	\item Replay attacks
+	\item Attacks against intermediaries: because of the hierarchic and centralized approch of DNS (there is only one root), the responsability of intermediaries is really high and a remote control by a thief should be bad (the thief controls the domain thus).
+History has proved that some of those intermediaries are not that trustworthy \cite{dan08} \cite{utk11}.
+\end{itemize}
+
+To make a conclusion about DANE, we have multiple points to point out:
+\begin{itemize}
+	\item Introducing a new register TLSA shall be very complicated, knowing that things are not moving that fast in DNS\footnote{Like it use to be with EDNS introduction few years ago.}.
+	\item The validation necessity end to end is quite disturbing knowing that there is no standart and deployable solution for wide use  about the last kilometer securisation.
+	\item DNSSEC is not possible behind a DNS proxy (which filter/break DNS answers to the point of creating a DoS for all TLS services.
+\end{itemize}
+
+Therefor, DANE is on a good way, particulary for communications between servers, where there is no (not much) constraints of security are (more) easily achievable.
 
@@
 
+\bibitem{dan08} Dancho Danchev. Hackers hijack dns records of high profile new zealand sites. http://m.zdnet.com/blog/security/hackers-hijack-dns-records-of-high-profile-new-zealand-sites/3185, 2008.
+
+\bibitem{utk11} Utkarsh. Domaining.com hacked. http://dotsutra.com/domain-names/industry-news/domaining-com-hacked/, 2011.
+
 \end{thebibliography}

### 2015-12-25 paper.tex
title Sovereign keys
Orthograph
Working on sovereign keys

@@
 % Define document title and author
-	\title{From a centralized certification systems to a more internet-like one}
+	\title{From centralized certification systems to more internet-like ones}
 	\author{Alexandre Kervadec
@@
 \end{itemize}
+
+\begin{figure}[!hbt]
+	\begin{center}
+	\includegraphics[width=\columnwidth]{pki.png}
+	\label{fig:pki}
+	\caption{Diagram of the X.509 and CAs process (from \cite{wang2015domain})}
+	\end{center}
+\end{figure}
 
@@
 
+\begin{figure}[!hbt]
+	\begin{center}
+	\includegraphics[width=\columnwidth]{danereq.png}
+	\label{fig:pki}
+	\caption{Process of certificate verification (from \cite{wang2015domain})}
+	\end{center}
+\end{figure}
+
 As we said before, DNS is not secured, so DNSEC has been deployed.
@@
 Therefor, DANE is on a good way, particulary for communications between servers, where there is no (not much) constraints of security are (more) easily achievable.
+
+\subsection{Sovereign keys}
+\label{sovkeys}
+
+Sovereign keys are an Electric Frontier Foundation (EFF) proposition.
+
+A sovereign key is a cryptographic pair of key where the public key is associated to a domain name (or a subdomain) to create a trust anchor\footnote{A trust anchor is an authoritative entity for which trust is assumed and not derived \cite{wallace2010}.} for itself.\\
+This association \textit{public key/domain name} is registered in a append-only data structure, kept on servers named ''\textit{timeline servers}''.
+There is a small number, around 10-20, of these \cite{eck2014}.\\
+The private part of the sovereign key (private key) allows to sign certificates of which the CN is the domain (or subdomain) name which it is associate with.\\
+During a handshake with a TLS server, the certificate is sent with its protected domain key signature, by the same mecanism than a TLS transaction based on PKIX.
+The client has then the possibility to recover the public key associated with that sovereign key from the timeline servers to check the key validity, that the domain is the same that the one in the current transaction and that it signs the receive certificate.
 
@@
 
+\bibitem{wallace2010} Wallace, C. Trust Anchor Format. No. RFC 5914. 2010.
+
+\bibitem{eck2014} P.
+Eckersley, ''Sovereign key cryptography for internet domains'', online, 2012, https://git.eff.org/?p=sovereign-keys.git;a=blob\_plain;f=sovereign-key-design.txt;hb=master, last retrieved on December 21, 2014.
+
+\bibitem{wang2015domain} Wang, Xinli, Yan Bai, and Lihui Hu.
+"Domain based certification and revocation." Proceedings of the 2015 International Conference on Security and Management, SAM.
+Vol.
+15.
+2015.
+
 \end{thebibliography}

### 2015-12-26 paper.tex
Sovereign keys -> DONE !

@@
 
-Sovereign keys are an Electric Frontier Foundation (EFF) proposition.
+Sovereign keys are an Electric Frontier Foundation (EFF\footnote{www.eff.org}) proposition.
 
@@
 The client has then the possibility to recover the public key associated with that sovereign key from the timeline servers to check the key validity, that the domain is the same that the one in the current transaction and that it signs the receive certificate.
+
+A sovereign key has a cycle of life than begins with the generation of a cryptographic pair of keys	by the domain administrator (who wants to use TLS server on his domain).
+This administrtor must at first, proove that he really is controlling his domain before seeing it in public data-structures.\\
+To do so, he must give his certificate (which the CN covers his domain name declared protected by the sovereign key) signed by an AC recognized by the \textit{timeline servers}\footnote{The list of AC's recognized by the timeline servers is public and must be up to date, by conception.} or through DANE (once the norm will by published and implemented).\\
+During the public key registering by the timeline servers, proofs of control on the domain are examined and added to the sovereign key metadata. Other informations are added sch as:
+\begin{itemize}
+	\item Expiration date.
+	\item Names (and ports) of TLS services uses and protected by the sovereign key.
+	\item A list (that may be empty) of domain names which their sovereign keys are granted to revok (in case of loose of control of the private part of the sovereign key) the entry being entered.
+\end{itemize}
+A data-structure of sovereign keys being a append-only list, once the sovereign key is entered, it cannot be removed.
+The domain administrator can only modify informations by pushing informations in that data-structure.
+To remove a sovereign key (in the case of losing the private key), the domain adminstrator need to use a domain name with a sovereign key granted to modify its own sovereign key (thus present in the list of his own sovereign key, see in the item list below) or wait for his sovereign key expiration date.
+
+This structure is useful because it protects from DNS or AC compromise of sovereign key by smurf attacks\footnote{Attacker able to give proof of domain control.}.
+It also need to be used with care, because an unadvise administrator may lose control of its own domain.\\
+
+The dark points of this systems are:
+\begin{itemize}
+	\item A service declared protected in a data-structure of a timeline server need to use a secure protocol (HTTPS instead of HTTP for example) to avoid downgrade attacks.
+It may be seen as an intrusion into the freedom of a service governance.
+	\item From loosing a private key may result a long term denial of service if no domain sovereign key is granted to modify the private key's sovereign key.
+	\item A compromized private key may lead to a total loose of control of the entire domain.
+	\item The use of ACs to certify the domain owner is the big problem of this system.
+\end{itemize}
 

### 2015-12-26 paper.tex
All done !

@@
 
-Going deeper with user validation : DN scheme based on X.500 Recomendation, but it is not completly defined, and will (in 1998) probably not be.
-Also, X509 certification depends on many others such as ISO, ANSI, ITU and IETF.
-Thus lead to a lack of harmonization.
-Plus, there is a big problem with CPS (Certification Practice Statements), that also can be seen such as flexibility (for pros), because each CA answer specific needs, so no harmonization again.
+There is a big problem with CPS (Certification Practice Statements), that can be seen such as flexibility (for pros), because each CA answer specific needs, so that is a sort of lack of harmonization.
 
-Also, regarrding to the law, CAs deny all leak of information from themselves.
+Also, regarding to the law, CAs deny all leak of information from themselves \cite{gerck1998overview}.
 
@@
 
+\section{Conclusion}
+\label{conclusion}
+
+Although the great need to replace the PKIX security model\footnote{X.509 and CAs}, none of those presented below may be valid, or well tested to be implemented in a large scale (world wide).
+
+Therefor, the best model may be the sovereign keys system associated with a DANE control of domain control certification, as soon as security problems will be fixed in DANE with DNSSEC.\\
+
+Otherwise, another point of view that need to be discussed, is the economic impact on such revolution: compagnies like VeriSign\footnote{www.verisign.com} or GeoTrust\footnote{www.geotrust.com} may face bankruptcy with the removal of all CAs.
+Knowing taht hey are big deciders in the internet activity, they detain a great lever on governements (because money leads capitalism), to stop or at least slow down the process of deployement of a new certification system.
+
+In general, to fit the Internet, experience has show us that we need to develop systems that are the most distributed as possible, like the internet paradigm.
+
+\pagebreak
+
 \nocite{*}
@@
 
+\bibitem{mf2011} Maury, Florian. "La fin annoncée des autorités de certification? Alternatives: TOFU, Convergence, CATA, Clés souveraines, DANE." Convergence 10 (2011): 509.
+
+\bibitem{correia2015dns} Correia, Miguel Medeiros, and Mustafa Tok. DNS-based Authentication of Named Entities (DANE). Technical report, Universidade do Porto, 2011–2012, 2015.
+
 \end{thebibliography}

### 2016-01-05 paper.tex
Review - version 1.1

@@
 \begin{abstract}
-This paper is a synthesis of a E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
+The current document is a study of a paper written by E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
 First, it gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
@@
 
-First sections talk about existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
-Then, further sections (\ref{dane} and \ref{certtrans}) deals with new, not implemented yet, technologies of digital certification which may solve current issues.
+X.509, CAs and PGP are the only technologies used to certify the authenticity of a user since the begining of the 90's.
+But those technologies appears to have issues, for example, Google ''became aware of unauthorized digital certificates for several Google domains.''\footnote{https://googleonlinesecurity.blogspot.fr/2013/12/further-improving-digital-certificate.html} This kind of threat led some researchers to find new certification systems.
+
+First sections shows existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
+Then, further sections (\ref{dane} and \ref{certtrans}) introduces new, not implemented yet, technologies of digital certification which may solve current issues.
 
@@
 
-\begin{figure}[!hbt]
+Lets get an example: \textit{user1} want to prove his identity to \textit{user2} with a CA (Figure 1).\\
+First, user1 needs to get a certificate signed by the CA, so he requests it to the CA with some informations about himself. The CA makes the certificate and sign it, then he send it back to user1.\\
+Now user2 wants to communicate with user1, so to proove his identity, user1 gives to user2 his certificate (forged by the CA).
+User2, in order to verify user1's identity, ask to the CA if the certificate is valid.
+Then the CA examine this certificate and answer to user2, the answer can be:
+\begin{itemize}
+	\item Yes, it is valid
+	\item No, it is not valid
+	\item Not sure about its validity
+\end{itemize}
+With that information, user2 can choose to trust or not user1.
+
+\begin{figure}[!h]
 	\begin{center}
-	\includegraphics[width=\columnwidth]{pki.png}
+		\includegraphics[scale=0.45]{pki.png}
+		\caption{Diagram of the X.509 and CAs process (from \cite{wang2015domain})}
 		\label{fig:pki}
-	\caption{Diagram of the X.509 and CAs process (from \cite{wang2015domain})}
 	\end{center}
@@
 Furthermore, SKIP does not support IP translation (such as NAT behind a firewall), that makes it quite useless.\\
-So that protocol imediatly looks unapropriate for our study. In other way, SKIP does not exist anymore, so it is pointless to go further in it study.
+So that protocol imediatly looks unapropriate for our study. In other way, SKIP does not exist anymore, so it is pointless to go further in its study.
 
@@
 
-To be used in a commercial way (for responsabilities purpose), X.509, CAs and PGP need a centralized certification control system.
-However, there is a struggle between this need of centralization and the internet architecture (that is decentralized).
-This show us the uselessness of centralized governement controls, since we have no central government on the internet which give a world governance.\\
-So the question emerging in there is: how can we (and do we need to) have a main control on an internet environement, to assure the goodness of datas and legitimacy of interlocutors?\\
-Some answers have been made since then (thus since the publication of the original article), answers that we are going to explain further.
+To be used in a commercial way, X.509, CAs and PGP need a centralized certification control system (a matter of responsibility).
+However, there is a struggle between this need of centralization and the decentralized internet architecture.
+It shows us the futility of certification authority for the whole world, since we have no centralized world governance or law.\\
+The main question is:\\
+\textit{How can we have a main control on an internet environement, to assure the security of the data and legitimacy of users/servers?}\\
+Some answers have been made since then (thus since the publication of the original article).
 
@@
 	\begin{center}
-	\includegraphics[width=\columnwidth]{danereq.png}
+	\includegraphics[scale=0.4]{danereq.png}
 	\label{fig:pki}
@@
 
-The dark points of this systems are:
+The drawbacks of this systems are:
 \begin{itemize}

### 2016-01-05 Presentation/main.tex
Initialization of presentation

\documentclass{bredelebeamer}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Certification systems]{From centralized certification systems to more internet-like ones}
% Titre du diaporama

\subtitle{}
% Sous-titre optionnel

\author{Alexandre Kervadec\inst{1}}
% La commande \inst{...} Permet d'afficher l' affiliation de l'intervenant.
% Si il y a plusieurs intervenants: Marcel Dupont\inst{1}, Roger Durand\inst{2}
% Il suffit alors d'ajouter un autre institut sur le modèle ci-dessous.

\institute[Université de Bordeaux]
{
  \inst{1}%
  Master 2 Informatique RSM
  }

\date{\today}
% Optionnel. La date, généralement celle du jour de la conférence

\subject{Overview of certifcation systems}
% C'est utilisé dans les métadonnes du PDF

\logo{
\includegraphics[scale=0.15]{images/logo.jpg}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Sommaire}
  \tableofcontents
  % possibilité d'ajouter l'option [pausesections]
\end{frame}

\section{Les systèmes en place}

\begin{frame}{Les systèmes en place}

\begin{description}
	\item [Certificats X.509 et CAs] (Certification Authority)
	\item [PGP] (Pretty Good Privacy)
	\item [SKIP] (Simple Key management for Internet Protocols)
\end{description}

\end{frame}

\begin{frame}{Questions}

	%\begin{figure}
	%	\centering
	%	\includegraphics[scale=0.27]{images/questions.jpg}
	%\end{figure}

\end{frame}

\end{document}

### 2016-01-07 paper.tex
Changed pki example naming

@@
 
-Lets get an example: \textit{user1} want to prove his identity to \textit{user2} with a CA (Figure 1).\\
-First, user1 needs to get a certificate signed by the CA, so he requests it to the CA with some informations about himself. The CA makes the certificate and sign it, then he send it back to user1.\\
-Now user2 wants to communicate with user1, so to proove his identity, user1 gives to user2 his certificate (forged by the CA).
-User2, in order to verify user1's identity, ask to the CA if the certificate is valid.
-Then the CA examine this certificate and answer to user2, the answer can be:
+Lets get an example: \textit{user1} want to prove his identity to \textit{Bob} with a CA (Alice) (Figure 1).\\
+First, Oscar needs to get a certificate signed by Alice, so he requests it to Alice with some informations about himself. Alice makes the certificate and signs it, then she sends it back to Oscar.\\
+Now Bob wants to communicate with Oscar, so to proove his identity, Oscar gives to Bob his certificate (forged by Alice).
+Bob, in order to verify Oscar's identity, ask to Alice if the certificate is valid.
+Then Alice examine this certificate and answer to Bob, the answer can be:
 \begin{itemize}
@@
 \end{itemize}
-With that information, user2 can choose to trust or not user1.
+With that information, Bob can choose to trust or not Oscar.
 

### 2016-01-07 paper.tex
Changed the wrong names in the PKI example | reorganized PGP part

@@
 
-Lets get an example: \textit{user1} want to prove his identity to \textit{Bob} with a CA (Alice) (Figure 1).\\
-First, Oscar needs to get a certificate signed by Alice, so he requests it to Alice with some informations about himself. Alice makes the certificate and signs it, then she sends it back to Oscar.\\
-Now Bob wants to communicate with Oscar, so to proove his identity, Oscar gives to Bob his certificate (forged by Alice).
-Bob, in order to verify Oscar's identity, ask to Alice if the certificate is valid.
-Then Alice examine this certificate and answer to Bob, the answer can be:
+Lets get an example: \textit{user1} want to prove his identity to \textit{Bob} with a CA (Jack) (Figure 1).\\
+First, Oscar needs to get a certificate signed by Jack, so he requests it to Jack with some informations about himself. Jack makes the certificate and signs it, then she sends it back to Oscar.\\
+Now Bob wants to communicate with Oscar, so to proove his identity, Oscar gives to Bob his certificate (forged by Jack).
+Bob, in order to verify Oscar's identity, ask to Jack if the certificate is valid.
+Then Jack examine this certificate and answer to Bob, the answer can be:
 \begin{itemize}
@@
 
+Thus, we can figure out that PGP breaks the traditional hierarchical trust architecture with this \textit{web-of-trust} approach.\\
+
 \begin{figure}[!hbt]
@@
 
-Thus, we can figure out that PGP breaks the traditional hierarchical trust architecture with this \textit{web-of-trust} approach.\\
-
-Let's get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}}: Bob want to exchange datas with Oscar, which he didn't knew before.
+Let's get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}} (Figure \ref{fig:web-of-trust}): Bob want to exchange datas with Oscar, which he didn't knew before.
 But Jack, a colleague of Bob's, signs Oscar's public-key certificate whch he knows is authentic.\\
@@
 This is illustrated in Figure \ref{fig:web-of-trust}.\\
-This is illustrated in Figure 1, where the dotted arrow indicates ''Bob trusts Jack as introducer'' and the solid arrow indicates ''Jack trusts Oscar’s public key validity''.
+The dotted arrow indicates ''Bob trusts Jack as introducer'' and the solid arrow indicates ''Jack trusts Oscar’s public key validity''.
 

### 2016-01-07 paper.tex
Orthograph review

@@
 The current document is a study of a paper written by E.Gerck's paper : \textit{"Overview of Certification systems: X.509, CA, PGP and SKIP"}\cite{gerck1998overview}, but not only.
-First, it gives an overview of (the mains) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thoughts on pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
+First, it gives an overview of (the main) certification systems which are X.509 and CAs, PGP, SKIP, DANE and the Certificate transparency by Google, with thought on the pros and cons of each system, on a technical point of view and about the government's stranglehold on data exchanges.
 \end{abstract}
@@
 
-X.509, CAs and PGP are the only technologies used to certify the authenticity of a user since the begining of the 90's.
-But those technologies appears to have issues, for example, Google ''became aware of unauthorized digital certificates for several Google domains.''\footnote{https://googleonlinesecurity.blogspot.fr/2013/12/further-improving-digital-certificate.html} This kind of threat led some researchers to find new certification systems.
+X.509, CAs and PGP are the only technologies used to certify the authenticity of a user since the beginning of the 90's.
+But those technologies appear to have issues, for example, Google ''became aware of unauthorized digital certificates for several Google domains.''\footnote{https://googleonlinesecurity.blogspot.fr/2013/12/further-improving-digital-certificate.html} This kind of threat led some researchers to find new certification systems.
 
-First sections shows existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
-Then, further sections (\ref{dane} and \ref{certtrans}) introduces new, not implemented yet, technologies of digital certification which may solve current issues.
+First sections show existing (sections \ref{x509} and \ref{pgp}) or extinguished (section \ref{skip}) certification systems.
+Then, further sections (\ref{dane} and \ref{certtrans}) introduce new, not implemented yet, technologies of digital certification which may solve current issues.
 
@@
 X.509 and CAs\cite{rfc3647} infrastructure is based on a directory method.\\
-With this kind of certification system, there are three different entities, which are :
+With this kind of certification system, there are three different entities:
 
@@
 	\item \textbf{CA : Certification Authority}, an entity that controls authentication services and management of digital certificates.
-It can be public (like banks with their clients), commercial (such as \textit{Verisign} which sells its services) or private (like a compagny department, for an internal purposes).
-	\item \textbf{Subscriber :} the entity which sends informations to the CA to add it to his certificate. This entity is one that need to be trusted by the \textit{user} entity.
-	\item \textbf{User :} ask informations about \textit{subscribers} to CA(s), it's central to the process, since the user party is relying on the informations given by \textit{CAs} and is thus at risk.
+It can be public (like banks with their clients), commercial (such as \textit{Verisign} which sells its services) or private (like a company department, for an internal purposes).
+	\item \textbf{Subscriber :} the entity which sends data to the CA to add it to his certificate. This entity is one that needs to be trusted by the \textit{user} entity.
+	\item \textbf{User :} ask information about \textit{subscribers} to CA(s), it is central to the process, since the user party is relying on the informations given by \textit{CAs} and is thus at risk.
 \end{enumerate}
 
-The way that certification system is the following:
+The way that certification system works is the following:
 \begin{itemize}
 	\item A subscriber gives informations to the CA, in order to create a digital certificate for itself in the CA directory.
-	\item The CA receive those informations and put it in a certificate.
-	\item A user (the public in general) receive a certificate from a subscriber.
-In order to authenticate this subscriber, the client ask to the CA if the transmitted certificate matches with the subscriber's identity.
-	\item The CA answer to the user to give him the precious informations about the subscriber's identity.
+	\item The CA receives the information and puts it in a certificate.
+	\item A user (the public in general) receives a certificate from a subscriber.
+In order to authenticate this subscriber, the client asks to the CA if the transmitted certificate matches with the subscriber's identity.
+	\item The CA answers to the user to give him the precious informations about the subscriber's identity.
 \end{itemize}
 
-Lets get an example: \textit{user1} want to prove his identity to \textit{Bob} with a CA (Jack) (Figure 1).\\
-First, Oscar needs to get a certificate signed by Jack, so he requests it to Jack with some informations about himself. Jack makes the certificate and signs it, then she sends it back to Oscar.\\
-Now Bob wants to communicate with Oscar, so to proove his identity, Oscar gives to Bob his certificate (forged by Jack).
-Bob, in order to verify Oscar's identity, ask to Jack if the certificate is valid.
-Then Jack examine this certificate and answer to Bob, the answer can be:
+Let us get an example: \textit{user1} wants to prove his identity to \textit{Bob} with a CA (Jack) (Figure 1).\\
+First, Oscar needs to get a certificate signed by Jack, so he requests it from Jack with some informations about himself. Jack makes the certificate and signs it, then he sends it back to Oscar.\\
+Now Bob wants to communicate with Oscar, so to prove his identity, Oscar gives to Bob his certificate (forged by Jack).
+In order to verify Oscar's identity, Bob asks to Jack if the certificate is valid.
+Then Jack examines this certificate and answers to Bob, the answer can be:
 \begin{itemize}
@@
 There is a not outwardly perceived entity in this system, it is the Naming Authority (DN).
-It provides to the CA a unique distinguished name (DN) which matches to the subscriber identity in the certificate.\\
+It provides to the CA a unique distinguished name (DN) which matches with the subscriber's identity in the certificate.\\
 Note that the CA can double in NA, but they provide two different services: the CA certificate refers to a name but does not denote it, the NA does.\\
@@
 	\item The content of a certificate needs to be discussed, as well as certificate revocation.
-For example, a subscriber can generate multiple DN for a same CA or the same DN to multiple CAs, so maybe, a better subscriber identificator need to be added.
+For example, a subscriber can generate multiple DN for a same CA or the same DN to multiple CAs, so maybe, a better subscriber identificator needs to be added.
 	\item Are the validation procedures for the certified data that is included in a certificate solid enough?
-Because each CA has its own self-defined rules (Certification Practice Statement\footnote{CPS}), which can by completly different from one CA to another.
+Because each CA has its own self-defined rules (Certification Practice Statement\footnote{CPS}), which can be completly different from one CA to another.
 \end{itemize}
 
-There is a big problem with CPS (Certification Practice Statements), that can be seen such as flexibility (for pros), because each CA answer specific needs, so that is a sort of lack of harmonization.
+There is a big problem with CPS (Certification Practice Statements), that can be noted such as flexibility (for pros), because each CA answer has specific needs, so that is a sort of lack of harmonization.
 
-Also, regarding to the law, CAs deny all leak of information from themselves \cite{gerck1998overview}.
+Also, regarding the law, CAs deny any leak of information from themselves \cite{gerck1998overview}.
 
@@
 
-PGP, created thanks to Phil Zimmermann researches. It has two parts: certification and encryption, following text only deals with certification.\\
+PGP, created thanks to Phil Zimmermann research. It has two parts: certification and encryption, the following text only deals with certification.\\
 Compared to X.509, PGP is more \textit{internet-like}, this because of its \textit{introducer-model} base.\\
-PGP depends on a chain of authentificators: the users themselves.
+PGP depends on a chain of authenticators: the users themselves.
 This chain forms a ring, but not in a closed way but in a mathematical way like a list or a \textit{web-of-trust}.
-You may not know everyone in this ring, but you can assume that you will know somebody who know this user.
-You can also assume that different rings can have some contact points to garantee the referrals.
+You may not knows everyone in this ring, but you can assume that you will know somebody who know this user.
+You can also assume that different rings can have some contact points to guarantee the referrals.
 
@@
 
-Let's get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}} (Figure \ref{fig:web-of-trust}): Bob want to exchange datas with Oscar, which he didn't knew before.
-But Jack, a colleague of Bob's, signs Oscar's public-key certificate whch he knows is authentic.\\
+Let us get an example\footnote{Example from ''The PGP trust model''\cite{caronni2000}} (Figure \ref{fig:web-of-trust}): Bob wants to exchange data with Oscar, whom he didn't know before.
+But Jack, a colleague of Bob's, signs Oscar's public-key certificate which he knows is authentic.\\
 Oscar then forwards his signed certificate to Bob who wishes to communicate with Oscar privately.\\
-Bob, who knows and truts Jack as an \textit{introducer}, find out, after verification, that Jack is among Oscar's certificate signer.
-Therefor, Bob can be confident that Oscar's public key is authentic.
+Bob, who knows and trusts Jack as an \textit{introducer}, finds out, after verification, that Jack is among Oscar's certificate signer.
+Therefore, Bob can be confident that Oscar's public key is authentic.
 
 However, had Bob not known or trusted Oscar's signers, including Jack, he would have been skeptical about the authenticity of Oscar's public-key.
-Oscar would have to find another introducer whom Bob trusts to sign Oscar's public-key certificate.
+Oscar would then have to find another introducer whom Bob trusts to sign Oscar's public-key certificate.
 This is illustrated in Figure \ref{fig:web-of-trust}.\\
@@
 Only the trustworthiness of public-key's validity is automatically evaluated by PGP.
-Introducer's worthiness is manually assigned by each user of to the public key, and exists only within each individual user's public-key ring\footnote{PGP allows users to have files representing multipke \textit{key rings} to store public or secret keys\cite{caronni2000}.}.
+The introducer's worthiness is manually assigned by each user of to the public key, and exists only within each individual user's public-key ring\footnote{PGP allows users to have files representing multipke \textit{key rings} to store public or secret keys\cite{caronni2000}.}.
 
-The big point is the use of such a tool in a world wide range: that is raisonnably not possible.
-We can assume to use it in a little group, where trustworthiness is effective, but in a more wide area, it is problematic to trust such a amount of people, or find trusted introducers.
+The big point is the use of such a tool in a world wide range: that is reasonably not possible.
+We can assume it can be used in a little group, where trustworthiness is effective, but in a wider area, it is problematic to trust such a number of people, or find trusted introducers.
 Furthermore, in a commercial use, there is no entity responsible when something goes wrong.
@@
 Furthermore, SKIP does not support IP translation (such as NAT behind a firewall), that makes it quite useless.\\
-So that protocol imediatly looks unapropriate for our study. In other way, SKIP does not exist anymore, so it is pointless to go further in its study.
+So that protocol imediatly looks inappropriate for our study. In an other way, SKIP does not exist anymore, so it is pointless to go further in its study.
 
@@
 The main question is:\\
-\textit{How can we have a main control on an internet environement, to assure the security of the data and legitimacy of users/servers?}\\
-Some answers have been made since then (thus since the publication of the original article).
+\textit{How can we have a main control over an internet environment, to assure the security of the data and legitimacy of users/servers?}\\
+Some answers have been provided since then (thus since the publication of the original article).
 
@@
 
-DANE is not implemented yet, but a IETF team is working on the standard.\\
-It has begin with a simple assert: why use a new trust infrastructure if we already have to use DNS to resolve a domain name?\\
+DANE is not implemented yet, but an IETF team is working on the standard.\\
+It has begun with a simple assertion: why use a new trust infrastructure if we already have to use DNS to resolve a domain name?\\
 From here, DNS must be securised, knowing that it is easily to set a MITM with DNS cache poisoning \cite{aa04} using the Kaminsky attack \cite{bor08}.
 That is the main purpose of DNSSEC: secure DNS.
-DANE allow to store keys in DNS.
+DANE allows to store keys in DNS.
 Thereby, \verb$www.example.org$ would be able to secure \verb$https://www.example.org$ without using a suspicious third part, just by putting its keys in the domain name that it controls.
@@
 	\item \textit{Type 0:} Alice has got a certificate delivered by Charlie and is really satisfied by it.
-But Alice is afraid that \textit{Oscar} (a bad AC) forge a false certificate for its website.
+But Alice is afraid that \textit{Oscar} (a bad AC) might forge a false certificate for his website.
 So Alice uses DANE to tell in DNS that Charlie and only Charlie is its AC.
-	\item \textit{Type 1:} Alice has got a certificate delivered by Charlie, but fear that Charlie give its certificate to bad people. So Alice uses DANE to publish in DNS the origin of its certificate.
-	\item \textit{Type 2:} that is the \textit{max scenario} in DANE. Alice doesn't trust ACs anymore, so it decides to emite it via DANE on DNS, that gives the entire responsability on DNS servers. 
+	\item \textit{Type 1:} Alice has got a certificate delivered by Charlie, but fears that Charlie might give its certificate to bad people.
+So Alice uses DANE to publish in DNS the origin of its certificate.
+	\item \textit{Type 2:} that is the \textit{max scenario} in DANE. Alice doesn't trust ACs anymore, so it decides to emit it via DANE on DNS, that gives the entire responsibility on DNS servers. 
 \end{itemize}
@@
 As we said before, DNS is not secured, so DNSEC has been deployed.
-DNSSEC is a protocol that allow to the DNS server (master) ti encrypt and sign its answer with a public key cryptographic system.
-It also permit to return a \verb$NXDOMAIN$ type (''No Such Domain'').\\
+DNSSEC is a protocol that allows to the DNS server (master) to encrypt and sign its answer with a public key cryptographic system.
+It also permits to return a \verb$NXDOMAIN$ type (''No Such Domain'').\\
 Nervertheless DNSSEC has some vulnaribilities such as:
@@
 	\item Replay attacks
-	\item Attacks against intermediaries: because of the hierarchic and centralized approch of DNS (there is only one root), the responsability of intermediaries is really high and a remote control by a thief should be bad (the thief controls the domain thus).
+	\item Attacks against intermediaries: because of the hierarchic and centralized approach of DNS (there is only one root), the responsibility of intermediaries is really high and a remote control by a thief should be damaging (the thief controls the domain thus).
 History has proved that some of those intermediaries are not that trustworthy \cite{dan08} \cite{utk11}.
@@
 	\item Introducing a new register TLSA shall be very complicated, knowing that things are not moving that fast in DNS\footnote{Like it use to be with EDNS introduction few years ago.}.
-	\item The validation necessity end to end is quite disturbing knowing that there is no standart and deployable solution for wide use  about the last kilometer securisation.
-	\item DNSSEC is not possible behind a DNS proxy (which filter/break DNS answers to the point of creating a DoS for all TLS services.
+	\item The end to end validation necessity end to end is quite disturbing knowing that there is no standard and deployable solution for wide use  about the last kilometer securisation.
+	\item DNSSEC is not possible behind a DNS proxy (which filters/breaks DNS answers to the point of creating a DoS for all TLS services.
 \end{itemize}
 
-Therefor, DANE is on a good way, particulary for communications between servers, where there is no (not much) constraints of security are (more) easily achievable.
+Therefore, DANE is on a good way, particularly for communications between servers, where there is no (not many) constraints of security are (more) easily achievable.
 
@@
 
-A sovereign key is a cryptographic pair of key where the public key is associated to a domain name (or a subdomain) to create a trust anchor\footnote{A trust anchor is an authoritative entity for which trust is assumed and not derived \cite{wallace2010}.} for itself.\\
+A sovereign key is a cryptographic pair of keys where the public key is associated to a domain name (or a subdomain) to create a trust anchor\footnote{A trust anchor is an authoritative entity for which trust is assumed and not derived \cite{wallace2010}.} for itself.\\
 This association \textit{public key/domain name} is registered in a append-only data structure, kept on servers named ''\textit{timeline servers}''.
 There is a small number, around 10-20, of these \cite{eck2014}.\\
-The private part of the sovereign key (private key) allows to sign certificates of which the CN is the domain (or subdomain) name which it is associate with.\\
-During a handshake with a TLS server, the certificate is sent with its protected domain key signature, by the same mecanism than a TLS transaction based on PKIX.
-The client has then the possibility to recover the public key associated with that sovereign key from the timeline servers to check the key validity, that the domain is the same that the one in the current transaction and that it signs the receive certificate.
+The private part of the sovereign key (private key) allows to sign certificates of which the CN is the domain (or subdomain) name which it is associated to.\\
+During a handshake with a TLS server, the certificate is sent with its protected domain key signature, by the same mechanism than as a TLS transaction based on PKIX.
+The client has then the possibility to recover the public key associated with that sovereign key from the timeline servers to check the key's validity, that the domain is the same as the one in the current transaction and that it signs the receive certificate.
 
-A sovereign key has a cycle of life than begins with the generation of a cryptographic pair of keys	by the domain administrator (who wants to use TLS server on his domain).
-This administrtor must at first, proove that he really is controlling his domain before seeing it in public data-structures.\\
-To do so, he must give his certificate (which the CN covers his domain name declared protected by the sovereign key) signed by an AC recognized by the \textit{timeline servers}\footnote{The list of AC's recognized by the timeline servers is public and must be up to date, by conception.} or through DANE (once the norm will by published and implemented).\\
-During the public key registering by the timeline servers, proofs of control on the domain are examined and added to the sovereign key metadata. Other informations are added sch as:
+A sovereign key has a cycle of life that begins with the generation of a cryptographic pair of keys	by the domain administrator (who wants to use TLS server on his domain).
+This administrator must at first, prove that he really is controlling his domain before seeing it in public data-structures.\\
+To do so, he must give his certificate (of which the CN covers his domain name declared protected by the sovereign key) signed by an AC recognized by the \textit{timeline servers}\footnote{The list of AC's recognized by the timeline servers is public and must be up to date, by conception.} or through DANE (once the standard has been published and implemented).\\
+During the public key registering by the timeline servers, proofs of control on the domain are examined and added to the sovereign key metadata. Other data are added such as:
 \begin{itemize}
@@
 	\item Names (and ports) of TLS services uses and protected by the sovereign key.
-	\item A list (that may be empty) of domain names which their sovereign keys are granted to revok (in case of loose of control of the private part of the sovereign key) the entry being entered.
+	\item A list (that may be empty) of domain names whose sovereign keys are granted to revoke (in case of loss of control of the private part of the sovereign key) the entry being entered.
 \end{itemize}
 A data-structure of sovereign keys being a append-only list, once the sovereign key is entered, it cannot be removed.
-The domain administrator can only modify informations by pushing informations in that data-structure.
-To remove a sovereign key (in the case of losing the private key), the domain adminstrator need to use a domain name with a sovereign key granted to modify its own sovereign key (thus present in the list of his own sovereign key, see in the item list below) or wait for his sovereign key expiration date.
+The domain administrator can only modify informations by pushing data in that data-structure.
+To remove a sovereign key (in the case the private key has been lost), the domain adminstrator needs to use a domain name with a sovereign key granted to modify its own sovereign key (thus present in the list of his own sovereign keys, see in the item list below) or wait for his sovereign key's expiration date.
 
 This structure is useful because it protects from DNS or AC compromise of sovereign key by smurf attacks\footnote{Attacker able to give proof of domain control.}.
-It also need to be used with care, because an unadvise administrator may lose control of its own domain.\\
+It also needs to be used with care, because an unadvised administrator may lose control of its own domain.\\
 
@@
 \begin{itemize}
-	\item A service declared protected in a data-structure of a timeline server need to use a secure protocol (HTTPS instead of HTTP for example) to avoid downgrade attacks.
+	\item A service declared protected in a data-structure of a timeline server needs to use a secure protocol (HTTPS instead of HTTP for example) to avoid downgrade attacks.
 It may be seen as an intrusion into the freedom of a service governance.
-	\item From loosing a private key may result a long term denial of service if no domain sovereign key is granted to modify the private key's sovereign key.
-	\item A compromized private key may lead to a total loose of control of the entire domain.
-	\item The use of ACs to certify the domain owner is the big problem of this system.
+	\item Loosing a private key may result in a long term denial of service if no domain sovereign key is granted to modify the private key's sovereign key.
+	\item A compromized private key may lead to a total loss of control of the entire domain.
+	\item The use of ACs to certify the domain owner is the big flaw of this system.
 \end{itemize}
@@
 
-This protocol has been developped by a Google team few years ago \cite{LL11}.
-It is still in developpment \cite{certtransp2013} so it is not implemented and deployed yet.
-It also has to be improve, like everythign in computing sciences, in the early years.\\
+This protocol has been developed by a Google team a few years ago \cite{LL11}.
+It is still under development \cite{certtransp2013} so it is not implemented and deployed yet.
+It also has to be improve, like everything in computing science, in the early years.\\
 CATA's purpose is a public list of all certificates that are emitted by CAs.
-Thus, clients are allowed to search in that public list, and verify that the received certificate is in that list, to be validate (if it is not in the list, it shall be rejected).\\
+Thus, clients are allowed to search in that public list, and verify that the received certificate is in that list, to be validated (if it is not in the list, it shall be rejected).\\
 In that way, a(conscientious) domain administrator can have access to that list and monitor all fraudulent certificates for his domain.\\
@@
 	\item How many lists should be created?
-	\item How does the revokation should work?
+	\item How should the revokation work?
 	\item What should happen if a certificate is no longer available (path unreachable)?
@@
 
-Although the great need to replace the PKIX security model\footnote{X.509 and CAs}, none of those presented below may be valid, or well tested to be implemented in a large scale (world wide).
+Despite the great need to replace the PKIX security model\footnote{X.509 and CAs}, none of those presented above may be valid, or well tested to be implemented on a large scale (world wide).
 
-Therefor, the best model may be the sovereign keys system associated with a DANE control of domain control certification, as soon as security problems will be fixed in DANE with DNSSEC.\\
+Therefore, the best model may be the sovereign keys system associated with the DANE control of a domain control certification, as soon as security problems are fixed in DANE with DNSSEC.\\
 
-Otherwise, another point of view that need to be discussed, is the economic impact on such revolution: compagnies like VeriSign\footnote{www.verisign.com} or GeoTrust\footnote{www.geotrust.com} may face bankruptcy with the removal of all CAs.
-Knowing taht hey are big deciders in the internet activity, they detain a great lever on governements (because money leads capitalism), to stop or at least slow down the process of deployement of a new certification system.
+Otherwise, another point of view that needs to be discussed, is the economic impact of such a revolution: companies like VeriSign\footnote{www.verisign.com} or GeoTrust\footnote{www.geotrust.com} may face bankruptcy with the removal of all CAs.
+Knowing that they are big deciders in the internet activity, they detain a great lever on governments (because money leads capitalism), to stop or at least slow down the deployment process of a new certification system.
 
-In general, to fit the Internet, experience has show us that we need to develop systems that are the most distributed as possible, like the internet paradigm.
+In general, to fit the Internet, experience has shown that we need to develop systems that are the most distributed possible, like the internet paradigm.
 

### 2016-01-17 Presentation/main.tex
finished presentation slides

@@
 \documentclass{bredelebeamer}
+
+\usepackage{tikz}
+\usepackage{verbatim}
+\usepackage{fancyvrb}
+\usepackage{listings}
 
@@
 
+\subsection{X.509 et CAs}
+
+\begin{frame}{X.509 et CAs}
+
+	On distingue 3 différentes entités :
+	\begin{description}
+		\item [Certification Authority\footnote{CA}] : entité qui contrôle l'authentification et la gestion des certificats digitaux
+		\item [Subscriber] : l'entité qui envoie des données au \textbf{CA} pour les ajouter à son certificat. C'est l'entité en qui l'\textbf{user} veut avoir confiance
+		\item [User] : demande des information au \textbf{CA} sur la validité du certificat que lui a envoyé un \textbf{subscriber}
+	\end{description}
+
+\end{frame}
+
+\begin{frame}{X.509 et CAs}
+
+	\begin{figure}
+		\centering
+		\includegraphics[scale=0.45]{./../pki.png}
+	\end{figure}
+
+\end{frame}
+
+\subsection{PGP}
+
+\begin{frame}{PGP}
+
+	\begin{figure}[!hbt]
+        \begin{center}
+
+                \begin{tikzpicture}[shorten >=1pt,->]
+                        \tikzstyle{vertex}=[circle,fill=black!25,minimum size=25pt,inner sep=0pt]
+
+                        \foreach \name/\angle/\text in {P-1/234/Alice, P-2/162/Bob,
+                                  P-3/90/Oscar, P-4/18/Jack, P-5/-54/Alex}
+                                \node[vertex,xshift=6cm,yshift=.5cm] (\name) at (\angle:2cm) {$\text$};
+
+                        \draw (P-1) -- (P-5);
+                        \draw[dashed,->] (P-2) -- (P-4);
+                        \draw (P-4) -- (P-3);
+                        \draw (P-3) -- (P-1);
+                        \draw (P-1) -- (P-2);
+                        \draw (P-2) -- (P-1);
+                        \draw (P-3) -- (P-5);
+                        \draw (P-5) -- (P-3);
+
+                \end{tikzpicture}
+                \caption{Communities of trust: Jack introduces Oscar to Bob's public-key certificate before Bob receives it.}
+                \label{fig:web-of-trust}
+        \end{center}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}{PGP}
+
+	PGP inclut un certificat à clefs publiques et une chaîne de confiance(un \textbf{introducer}).\\
+	\hfill \\
+	Les niveaux de confiance pour un certificat sont :
+	\begin{description}
+		\item [Undefined] : on ne peut pas dire si la cle publique est valide ou pas
+		\item [Marginal] : la clef publique \textbf{peut} être valide, mais on ne peut pas en être sûr
+		\item [Complete] : on est sûr que la clef publique est valide
+	\end{description}
+
+\end{frame}
+
+\begin{frame}{PGP}
+
+	Les niveaux de confiance pour une chaîne de confiance (\textbf{introducer}) sont :
+	\begin{description}
+		\item [Full] : on peut faire totalement confiance à la clef publique pour introduire une nouvelle clef publique 
+		\item [Marginal] : la clef publique \textbf{peut} introduire une nouvelle clef publique, mais on ne peut pas en être sûr
+		\item [Untrustworthy] : on ne doit pas faire confiance à cette clef publique pour en introduire une nouvelle
+	\end{description}
+
+\end{frame}
+
+\subsection{SKIP}
+
+\begin{frame}{SKIP}
+
+	\textbf{SKIP} (Simple Key-Management for Internet Protocol) implémente une chaîne d'authentificateurs de noeuds, où chaque noeud stocke ses informations dans une sorte de \textbf{directory service}\footnote{annuaire}.\\
+	\hfill\\
+	Etant donné que SKIP ne supporte pas la translation d'adresse (NAT par exemple), ce système est inutile.
+
+\end{frame}
+
+\section{Les systèmes en cours de développement}
+
+\begin{frame}{Les systèmes en cours de développement}
+
+	\begin{description}
+		\item [DANE] (DNS-Based Authentication of Named Entities)
+		\item [Sovereign keys] Les clefs souveraines par l'EFF\footnote{Electric Frontier Foundation}
+		\item [CATA] (Certificate Authority Transparency and Auditability)
+	\end{description}
+
+\end{frame}
+
+\subsection{DANE}
+
+\begin{frame}{DANE}
+
+	Ce système est arrivé avec une simple question : \textit{Pourquoi utiliser une nouvelle infrastructure de confiance alors que nous utilisons déjà DNS pour résoudre les noms de domaine ?}\\
+	\hfill\\
+	DANE utilise un nouveau champ dans la requête DNS\footnote{Sécurisé avec DNSEC} : le champ \textbf{TLSA}.
+\begin{center}
+\texttt{\_443.\_tcp.www.example.com. IN TLSA (\\
+	~~~~~0 0 1 d2abde240d7cd3ee6b4b28c54df0\&34b9\\
+	~~~~~7983a1d16e8a410e4561cb106618e971 )}
+\end{center}
+
+\end{frame}
+
+\begin{frame}{DANE}
+
+	\begin{figure}
+		\centering
+		\includegraphics[scale=0.35]{./../danereq.png}
+		\caption{Fonctionnement de DANE}
+	\end{figure}
+
+\end{frame}
+
+\subsection{Sovereign keys}
+
+\begin{frame}{Sovereign keys}
+
+	Le système de clefs souveraines s'appuie sur des \textbf{timeline servers}, sui seront en nombre restreint\footnote{Environ une vingtaine tout au plus}, gérant une liste publique dans laquelle on ne peut que faire des ajouts.\\
+
+	Des clefs souveraines sont un couple de clefs cryptographiques où la clef publique est associée à un nom de domaine.
+Cette association est enregistrée dans les \textbf{timeline servers} cités précédemment.\\
+	\hfill\\
+	Cycle de vie des clefs souverraines :
+	\begin{enumerate}
+		\item Génération de la pair de clefs 
+		\item Preuve de contrôle du domaine concerné (en utilisant PKI ou DANE)
+		\item Ajout de la clef publique associée au nom de domaine dans les timeline servers
+	\end{enumerate}
+
+\end{frame}
+
+\subsection{CATA}
+
+\begin{frame}{CATA}
+
+	Ce système se résume en une liste publique de certificats. Ainsi, chaque administrateur peut vérifier qu'il n'y a pas de fraude.\\
+	Ceci dit, beaucoup de questions restent à résoudre :
+	\begin{itemize}
+		\item Qui doit gérer ces listes ?
+		\item Combien de listes doivent être créées ?
+		\item Comment la révocation doit fonctionner ?
+		\item Qu'arrive-t-il si un certificat n'est plus disponible ?
+	\end{itemize}
+
+\end{frame}
+
 \begin{frame}{Questions}
-
-	%\begin{figure}
-	%	\centering
-	%	\includegraphics[scale=0.27]{images/questions.jpg}
-	%\end{figure}
+\begin{center}
+	Avez-vous des questions ?
+\end{center}
 

### 2016-01-29 Presentation/main.tex
Orthograph correction

@@
 
-	Le système de clefs souveraines s'appuie sur des \textbf{timeline servers}, sui seront en nombre restreint\footnote{Environ une vingtaine tout au plus}, gérant une liste publique dans laquelle on ne peut que faire des ajouts.\\
+	Le système de clefs souveraines s'appuie sur des \textbf{timeline servers}, qui seront en nombre restreint\footnote{Environ une vingtaine tout au plus}, gérant une liste publique dans laquelle on ne peut que faire des ajouts.\\
 
